D3064R0: OOTA Execution on Real Implementations is Provably Vacuous

### 2023-12-17 ootavacuous.tex
Fill out text for "Additional Litmus Tests"
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

@@

@@

@@

@@

>>>>

This appendix evaluates a number of OOTA-related litmus tests.
The evaluations are carried out using the \co{herd7} tool and also
using manual analysis.
The manual analysis uses the informal rule that a semantic dependency
exists between a given relaxed atomic load and store on any execution
where a different value loaded would change either the address or the
value of that store.
Producing a precise formal statement of this rule is quite challenging.

Appendix~\ref{app:Litmus Tests from “Causality Test Cases"}
evaluates the obligatory litmus tests from the JMM Causality Test
Cases, and
Appendix~\ref{app:Additional Litmus Tests}
evaluates some additional litmus tests that affected our thinking or
otherwise gave us pause.

@@

These tests might be two decades old, but they are still quite
relevant.

@@

These litmus tests helped illuminate important aspects of the problem
of defining sdep and identifying OOTA cycles.

@@

All processes's loads' return values affect the value stored, so all
processes have semantic dependencies between their respective relaxed
atomic loads and relaxed atomic stores.
Therefore, an OOTA cycle cannot form in a correct real-world C++
implementation.

@@

Figure~\ref{lst:Multi-Cycle Litmus Test}
shows an expansion to four threads of Causality Test Case 4, which is shown in
Listing~\ref{lst:Causality Test Case 4}
and discussed in
Appendix~\ref{app:Causality Test Case 4}.
The \co{herd7} tool shows multiple different possible OOTA cycles,
as expected.
However, the value returned by each thread's load directly affects the
value stored, so there is a semantic dependency in each thread.
Therefore, none of these OOTA cycles can happen in a correct real-world
C++ implementation.
